old snippets

### 2025-05-29 manuscript/masked_manuscript.qmd
fix repo

We tested a total of `r exclusions$sample$finish` participants from the Dutch Longitudinal Internet Studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
The LISS panel is an invitation-only, representative sample of the Dutch population consisting of approximately 7,500 individuals across 5,000 households.
LISS participants complete a yearly core battery of questionnaires about various domains of life, including one's financial situation over the past year.
In addition, LISS participants have the option to participate in further monthly studies on different topics.
We estimated power based on @kretzschmar_2019.
With $\alpha$ = 0.05 and assuming moderate reliability of our cognitive measures, we would have > 90% power to detect small effect sizes ($\beta$ = 0.1) with a sample size ranging between N = 730 and N = 980.
Therefore, we aimed for a total sample size of N = 1,000.

The current study took place between May and August 2024.
Participants were able to complete the cognitive tasks during two or more sessions to increase participation rates.
People were eligible for the study if they were between 18 and 55 years old, and had agreed to linking their LISS data to government microdata (not relevant here, but for a different study).
We excluded participants who did not have data on any cognitive task.
The final sample after exclusions consisted of `r exclusions$sample$final` participants (see Table 1).

@@

@@

```{r}
#| tab.id: table2
#| results: markup
table2 |> 
  flextable() |> 
  flextable::set_header_labels(Variable = "") |> 
  flextable::autofit() |> 
  border_remove() %>% 
 # border(i = 1, border.top = fp_border_default(), part = "header") %>% 
  border(i = 1, border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 18, border.bottom = fp_border_default(), part = "body") %>% 
  add_header_row(
    values = " ",
    colwidths = 13
  ) |> 
  flextable::compose(
    i = 1, j = 1, 
    as_paragraph(as_b("Table 2. "), "Spearman correlations between the main independent variables and covariates."),
    part = "header"
  ) |> 
  add_footer_row(
    values = " ",
    colwidths = 13
  ) |> 
  autofit()
```

@@

@@

The set of potential confounders consisted of (1) age [@salthouse_2016; @salthouse_2019; @starns_2010]; (2) education [@hofmarcher_2021]; (3) sex [@ning_2023]; (4) childhood adversity exposure (material deprivation and threat combined) [@bos_2009; @goodman_2019]; and (5) potential causal relations between recent material deprivation and recent threat [@bywaters_2016; @lacey_2022; @ning_2023].
We made these decisions based on previous literature, or, in cases of doubt, by statistically testing support for specific relations using data from a previous LISS study [@vermeent_2024c].
Specifically, following our preregistration, we used the *dagitty* R package [@ankan_2021; @textor_2016] to test whether independencies between specific variables implied by the Directed Acyclic Graph (conditional on the other causal pathways) were supported by these data (see the preregistration for more details).

Figure 1 depicts the final preregistered Directed Acyclic Graph for material deprivation and threat exposure in adulthood.
For threat exposure in adulthood, our statistical model controlled for age, sex, childhood adversity, and recent material deprivation.
We included material deprivation in adulthood based on studies showing that material deprivation tends to precede exposure to adversities such as threat [@bywaters_2016; @lacey_2022; @ning_2023].
We did not control for education given a lack of theoretical and statistical support for this effect (see preregistration).
To explore the effect of our assumption that material deprivation tends to precede threat, we also present a secondary model excluding material deprivation in adulthood as a confound.
As preregistered, we will not base our main conclusions on this secondary model.

For material deprivation in adulthood, our primary statistical model controlled for age, education, sex, and childhood adversity exposure based on previous studies (see above).
We did not control for threat exposure in adulthood, as, following the Directed Acyclic Graph, we assume that threat exposure mediates the effect of recent material deprivation on cognitive processes.
If true, including threat exposure in the model would make the association between material deprivation and cognitive processes an indirect effect, rather than a total effect (which was our estimand).
To explore the effect of our assumption that threat mediates the effect of material deprivation on cognitive processes, we also present a secondary model that includes threat in adulthood as a confound.
As preregistered, we will not base our main conclusions on this secondary model.

@@

We obtained ethical approval from the Ethics Review Board of the Faculty of Social & Behavioral Sciences of Utrecht University (FETC20-490) and the Ethics committee for research in the Sciences and Life Sciences of the University of Amsterdam (FNWI-41_2023).
The study was implemented on the LISS platform, and participants could only complete the study on a laptop or desktop PC.
Participants started with the six cognitive tasks, in randomized order.
After each task, they indicated their state anxiety and the level of environmental noise.
Then, participants had the option to either continue to the next task or complete the other tasks at a later point in time. 
`r round(n_sessions)` % of participants completed all tasks in a single session, and `r 100-round(n_sessions)` % of participants took two or more sessions.
Jointly, the cognitive tasks took around 25 minutes to complete.
After finishing all cognitive tasks, participants completed questionnaires about neighborhood threat and material deprivation during childhood and, adulthood exposure to crime victimization in the past two years.
They also completed questionnaires on impulsivity, self-control, and future orientation, which are not considered here because they fall outside of the scope of the present study.
As some participants had already completed all of these questionnaires in a different data collection (about six months prior to the current study), we only presented the questionnaires to new participants.
Finally, all participants answered a few standard LISS evaluation questions about the study, giving them an opportunity to provide written feedback.

@@

Each DDM parameter sub-model was a bi-factor model including the parameter estimates of all tasks as manifest variables.
We estimated a general factor accounting for variance in all manifest variables.
A second ability-specific inhibition factor loaded on the parameter estimates of incongruent trials of the Flanker and Simon tasks.
A third specific attention shifting factor loaded on the parameter estimates of switch trials of the Color-shape, Global-local, and Animacy-size tasks.
These two factors were allowed to covary.
We also estimated covariances between the conditions of each task.
For each sub-model, we compared the fit of this initial model to a version with a common EF factor loading on incongruent/switch trials of all tasks.
We deemed the second model a better fit when we observed a significant chi squared change test and an AIC value difference \> 10.

For the drift rate model, we interpreted the general factor as basic speed of processing, and the ability-specific factors as reflecting  inhibition and attention-shifting ability (or common EF in the case of the second model).
For the boundary separation and non-decision time models, we interpreted the general factor as general response caution/speed of non-decision processes, and the ability-specific factors as reflecting response caution/speed of non-decision processes specific to conflict trials.

@@

Finally, we constructed two versions of the final model, one to estimate the association between material deprivation in adulthood and the outcome measures, and one to estimate the association between threat exposure in adulthood and the outcome measures.
In both models, we regressed the adversity measure on each latent factor, together with the control variables (see the section on confounds for more details).
We report indirect effects of control variables in the supplemental materials (section 5).

@@

As preregistered, we started with optimizing fit in structural equation models for each DDM parameter separately before combining them all into a single model.
For drift rates, we selected the model containing a general processing speed factor (loading on all drift rates) and a common EF factor (loading only on drift rates of incongruent and switch conditions).
An alternative model with a separate inhibition and attention shifting factor did not converge.
The model with a common EF factor provided a good fit to the data (CFI = `r fit_v2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_v2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_v2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_v2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
However, the loadings of the common EF factor were small and in opposing directions, and the latent factor's residual variance was non-significant.
Therefore, we did not estimate associations between adversity and common EF.

@@

```{r}
#| label: Figure3
#| dpi: 600
#| out-width: 12in
#| fig-cap: | 
#|   **Figure 3.** Final measurement model of the Drift Diffusion parameters. Rectangles depict manifest parameter estimates, elipses depict latent factors, unidirectional arrows depict factor loadings, and bidirectional arrows depict correlations. Latent factors with non-significant residual variances are greyed out. *a* = manifest boundary separation, AS = Attention-shifting task, CS = Color-shape task, GL = Global-local task, Fl = Flanker task, Sm = Simon task, *t* = manifest non-decision time, *v* = manifest drift rate.
knitr::include_graphics("figures/fig3.png")
```

@@

```{r}
#| tab.id: table3
#| results: markup

table3 |> 
  add_header_row(
    values = " ",
    colwidths = 4
  ) |> # Add a new header row on top. We can use this new row to add the title
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table 3. "), "Fit statistics of all preregistered structural equation models."),
    part = "header"
  ) |>
  border_remove() |> 
  border(i = 1, border.bottom = fp_border_default(), part = "header") |>  
  border(i = 2, border.top = fp_border_default(), part = "header") |> 
  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 15, border.bottom = fp_border_default(), part = "body") |> 
  add_footer_row(
    values = " ",
    colwidths = 4
  ) |> 
  add_footer_row(
    values = " ",
    colwidths = 4
  ) |>  
  flextable::compose(
    i = 1, j = 1, 
    as_paragraph(as_i("Note: "), "for each DDM parameter, measurement model 1 contained a task-general latent factor as well as separate latent factors for inhibition and shifting. Measurement model 2 contained a task-general latent factor as well as a common EF factor. Measurement model 3 contained only a task-general latent factor."), 
    part = "footer"
  ) |> 
  autofit()

@@

Our primary analyses examined the association of exposure to material deprivation and threat in adulthood with task-general processing speed, task-general response caution, and task-general speed of stimulus encoding and/or response execution.
As the structural equation model analyses reported above showed that EF-specific factors had non-significant variances, we retained them in the model but did not include them as dependent variables in subsequent analyses.

@@

Exposure to material deprivation in adulthood was not associated with either general processing speed ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), task-general response caution ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
In addition, none of the effects fell within the region of practical equivalence (all *p*s ≥ `r fit_dep1_eqtests |> pull(eq_pvalue) |> min() |> format_p(pvalue = _, operator = FALSE)`).
The effects in the secondary model (including expossure to threat in adulthood as a confounder) were comparable.

@@

```{r}
#| label: Figure4
#| fig-width: 6.5
#| fig-height: 5.5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure 4.** Standardized (preregistered) associations between exposure to adversity in adulthood with task-general processing speed, response caution, and non-decision time. The left panel depicts effects of material deprivation and the right panel depicts effects of threat. The top row depicts effects of the primary statistical model (following from our main Directed Acyclic Graph), and the bottom row depicts effects of the secondary statistical model (see Figure 1). For material deprivation in adulthood, the primary statistical model included age, education, and childhood adversity as confounders, whereas the secondary model additionally included threat in adulthood as a confounder. For threat in adulthood, the primary statistial model included age, childhood adversity, and material deprivation in adulthood, whereas the secondary statistical model did not include material deprivation in adulthood as a confounder. The gray area reflects the preregistered area of practical equivalence. Effects depicted with a solid point are practically equivalent, and effects depicted with a hollow point are not practically equivalent. Standard errors represent 95% confidence intervals. Statistical significance (tested against zero) is indicated with asterisks; * *p* < .05, ** *p* < .01, *** *p* < .001.
fig4
```

@@

In the first set of non-preregistered analyses, we analyzed how childhood exposure to threat and material deprivation are associated with task-general and ability-specific cognitive processes.
The rational for these analyses is that we observed significant indirect associations between childhood adversity and general speed of processing in our preregistered analyses.
Here, we analyzed *direct* effects of childhood threat and material deprivation (i.e., without including exposure to adversity in adulthood).
We controlled for sex (but not age and education) and in the case of threat we also controlled for material deprivation, but not the other way around (similarly to the preregistered analyses).

@@

```{r}
#| label: Figure5
#| fig-width: 6.5
#| fig-height: 3
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure 5.** Standardized (non-preregistered) associations between childhood exposure to adversity with task-general processing speed, response caution, and non-decision time. The left panel depicts effects of material deprivation and the right panel depicts effects of threat. The gray area reflects the preregistered area of practical equivalence. Effects depicted with a solid point are practically equivalent, and effects depicted with a hollow point are not practically equivalent. Standard errors represent 95% confidence intervals. Statistical significance (tested against zero) is indicated with asterisks; * *p* < .05, ** *p* < .01, *** *p* < .001.
fig5
```

@@

In the second set of non-preregistered analyses, we analyzed how adversity exposure during childhood and adulthood were associated with task-specific residual variances.
The rationale for these analyses was that while we did not find coherent latent EF factors, all tasks showed significant residual variance after accounting for general speed of processing.
This implies that, after accounting for general speed of processing, performance is further determined by unique features of individual tasks (i.e., method variance), which may be associated with adversity exposure. 
For this analysis, we were mostly interested in investigating the extent to which task-specific effects reflect EF abilities or task-specific processing, and therefore only focused on drift rates.

@@

Figure 6 shows latent correlations between task-specific drift rates after accounting for task-general processing speed.
The drift rates of repeat and switch conditions within each attention shifting task correlated moderately to strongly, while all correlations between tasks were low.
This suggests that task-specific drift rates of attention shifting tasks captured features of the tasks that were not shared between tasks.
In contrast, all correlations between task-specific drift rates of inhibition tasks were low, even among drift rates of the same task.
Thus, task-specific drift rate did provide consistent measures of specific EF abilities that are shared across tasks.

@@

We investigated associations between exposure to two types of adversity in adulthood---material deprivation and threat---with inhibition and attention shifting ability.
Participants completed two inhibition tasks, three attention shifting tasks, and one basic processing speed task.
First, we used DDM to separate raw performance into three distinct cognitive processes: (1) speed of evidence accumulation (drift rate), (2) response caution (boundary separation), and (3) speed of stimulus encoding and response execution (non-decision time).
Finally, we used structural equation modeling to separate variance in each cognitive process into a task-general factor (shared across all tasks) and ability-specific factors.

@@

People with more exposure to threat in adulthood---but not material deprivation---showed lower general processing speed.
This is consistent with predictions from deficit frameworks [@sheridan_2014; @tucker_drob_2013].
After accounting for general processing speed, there was no remaining variance that could be attributed to either inhibition ability or attention shifting ability.
Additionally, the associations between material deprivation and threat in adulthood with both response caution and non-decision time were neither significantly different from zero, nor practically equivalent to zero.
In other words, from these data, we were neither able to conclude that people with more exposure to adversity differed in their level of response caution and non-decision time, nor that these processes were similar to people with lower levels of adversity exposure.

We found that variance in EF task performance was mostly explained by a task-general factor, not by ability-specific factors.
This finding aligns with previous research applying DDM to EF tasks. 
Three recent studies converged on the similar conclusion that performance across different EF tasks is best explained by a task-general drift rate factor [@frischkorn_2019; @hedge_2021; @loffler_2024; @weigard_2021].
One study found that the task-general drift rate factor fully reflected basic processing speed, rather than common EF processes [@loffler_2024].
Similarly, another study including several inhibition tasks found that shared variance was mostly explained by basic processing speed and strategy differences, not by inhibition ability [@hedge_2021].
It is therefore not surprising that we were unable to estimate ability-specific EF factors after accounting for general processing.
These issues tie into several other criticisms of common EF tasks in the cognitive psychology literature---such as low test-retest reliability---which should be considered when using these tasks to test associations with adversity [@vonbastian_2020].

>>>>

We tested a total of `r exclusions$sample$finish` participants from the Dutch Longitudinal Internet Studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
The LISS panel is an invitation-only, representative sample of the Dutch population consisting of approximately 7,500 individuals across 5,000 households.
LISS participants complete a yearly core battery of questionnaires about various domains of life, including one's financial situation over the past year.
In addition, LISS participants have the option to participate in further monthly studies on different topics.
Each month, participants have the opportunity to update their demographic information (e.g., sex, education level, household composition).
We estimated power based on @kretzschmar_2019.
With $\alpha$ = 0.05 and assuming moderate reliability of our cognitive measures, we would have > 90% power to detect small effect sizes ($\beta$ = 0.1) with a sample size ranging between N = 730 and N = 980.
Therefore, we aimed for a total sample size of N = 1,000.

The current study took place between May and August 2024.
Participants were able to complete the cognitive tasks during two or more sessions to increase participation rates.
People were eligible for the study if they were between 16 and 55 years old, and had agreed to linking their LISS data to government microdata (not relevant here, but for a different study).
We excluded participants who did not have data on any cognitive task.
The final sample after exclusions consisted of `r exclusions$sample$final` participants (see Table 1).

@@

### Childhood adversity

Participants reported their exposure to childhood deprivation (prior to age 18) in the new data collection using four items: (1) "My family had enough money to afford the kind of home we all needed"; (2) "My family had enough money to afford the kind of clothing we all needed"; (3) "My family had enough money to afford the kind of food that we all needed"; (4) "My family had enough money to afford the kind of medical care that we all needed", on a scale from 1 (completely not agree) to 7 (completely agree). 
This measure differed from those used to measure exposure to deprivation in adulthood. 
In addition, participants reported their exposure to childhood threat (prior to age 18) in the new data collection using the Neighborhood Violence Scale ("Crime was common in the neighborhood where I grew up") , on a scale from 1 (completely not agree) to 7 (completely agree) [@frankenhuis_2018; @frankenhuis_2020]. 
This measure was also used to measure threat exposure in adulthood.

When used as covariates in the main analyses (see Confounds section), the measures of childhood deprivation and childhood threat were averaged together into a single measure of childhood adversity. 
When used as main predictors in the exploratory analyses that we conducted, they were treated separately.

@@

```{r}
#| tab.id: table2
#| results: markup
table2 |> 
  flextable() |> 
  flextable::set_header_labels(Variable = "") |> 
  flextable::autofit() |> 
  border_remove() %>% 
 # border(i = 1, border.top = fp_border_default(), part = "header") %>% 
  border(i = 1, border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 20, border.bottom = fp_border_default(), part = "body") %>% 
  add_header_row(
    values = " ",
    colwidths = 15
  ) |> 
  flextable::compose(
    i = 1, j = 1, 
    as_paragraph(as_b("Table 2. "), "Spearman correlations between the main independent variables and covariates."),
    part = "header"
  ) |> 
  add_footer_row(
    values = " ",
    colwidths = 15
  ) |> 
  autofit()
```

@@

```{r}
table3 |> 
  mutate(empty = "") |> 
  select(task, condition, rt_m, rt_range, empty, ac_m, ac_range) |> 
  flextable() |> 
  set_header_labels(task = "Task", condition = "Condition", rt_m = "Mean (SD)", rt_range = "Range", empty = "", ac_m = "Mean (SD)", ac_range = "Range") |> 
  add_header_row(values = c("", "", "Response times", "", "", "Accuracy", "")) |> 
  merge_at(j = c(3:4), i = 1, part = "header") |> 
  merge_at(j = c(6:7), i = 1, part = "header") |>

  add_header_row(values = " ",colwidths = 7) |> 
  flextable::compose(
    i = 1, j = 1, 
    as_paragraph(as_b("Table 3. "), "Descriptive statistics of response times and accuracy on cognitive tasks."),
    part = "header"
  ) |> 
  align(i = 2, align = "center", part = "header") |> 
  border_remove() %>% 
  border(i = 2, border.top = fp_border_default(), part = "header") |> 
  border(i = 2, j = c(3,6), border.bottom = fp_border_default(), part = "header") |> 
  border(i = 3, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 11, border.bottom = fp_border_default(), part = "body") |>  
  autofit()
```

@@

The set of potential confounders consisted of (1) age [@salthouse_2016; @salthouse_2019; @starns_2010]; (2) education [@hofmarcher_2021]; (3) sex [@ning_2023]; (4) childhood adversity exposure (material deprivation and threat combined) [@bos_2009; @goodman_2019]; and (5) potential causal relations between recent material deprivation and recent threat [@bywaters_2016; @lacey_2022; @ning_2023].
We made these decisions based on previous literature, or, in cases of doubt, by statistically testing support for specific relations using data from a previous LISS study [@vermeent_2025].
Specifically, following our preregistration, we used the *dagitty* R package [@ankan_2021; @textor_2016] to test whether independencies between specific variables implied by the Directed Acyclic Graph (conditional on the other causal pathways) were supported by these data (see the preregistration for more details).

Figure 2 depicts the final preregistered Directed Acyclic Graph for material deprivation and threat exposure in adulthood.
For threat exposure in adulthood, our statistical model controlled for age, sex, childhood adversity, and recent material deprivation.
We included material deprivation in adulthood based on studies showing that material deprivation tends to precede exposure to adversities such as threat [@bywaters_2016; @lacey_2022; @ning_2023].
We did not control for education given a lack of theoretical and statistical support for this effect (see preregistration).
To explore the effect of our assumption that material deprivation tends to precede threat, we also present a secondary model excluding material deprivation in adulthood as a confound.
As preregistered, we will not base our main conclusions on this secondary model.

For material deprivation in adulthood, our primary statistical model controlled for age, education, sex, and childhood adversity exposure based on previous studies (see above).
We did not control for threat exposure in adulthood, as, following the Directed Acyclic Graph, we assume that threat exposure mediates the effect of recent material deprivation on cognitive processes.
If such mediation does occur, including threat exposure in the model would make the association between material deprivation and cognitive processes an indirect effect, rather than a total effect (which was our estimand).
To explore the effect of our assumption that threat mediates the effect of material deprivation on cognitive processes, we also present a secondary model that includes threat in adulthood as a confound.
As preregistered, we will not base our main conclusions on this secondary model.

@@

We obtained ethical approval from the Ethics Review Board of the Faculty of Social & Behavioral Sciences of Utrecht University (FETC20-490) and the Ethics committee for research in the Sciences and Life Sciences of the University of Amsterdam (FNWI-41_2023).
The study was implemented on the LISS platform, and participants could only complete the study on a laptop or desktop PC.
Participants started with the six cognitive tasks, in randomized order.
After each task, they indicated their state anxiety and the level of environmental noise.
Then, participants had the option to either continue to the next task or complete the other tasks at a later point in time. 
`r n_sessions |> filter(nomem_encr %in% exclusions$sample$ids) |> summarise(x = sum(n_sessions == 1)/n()*100) |> round(1) |> pull(x)`% of participants completed all tasks in a single session, and `r n_sessions |> filter(nomem_encr %in% exclusions$sample$ids) |> summarise(x = sum(n_sessions != 1)/n()*100) |> round(1) |> pull(x)`% of participants took two or more sessions.
Participants who took two or more sessions did not process information differently in the first session compared to later sessions (based on average standardized drift rates), *t*(`r session_effect$df_error`) = `r session_effect$t`, *p* `r format_p(session_effect$p)`.

Jointly, the cognitive tasks took around 25 minutes to complete.
After finishing all cognitive tasks, participants completed questionnaires about neighborhood threat and material deprivation during childhood and, adulthood exposure to crime victimization in the past two years.
They also completed questionnaires on impulsivity, self-control, and future orientation, which are not considered here because they fall outside of the scope of the present study.
As some participants had already completed all of these questionnaires in a different data collection (about six months prior to the current study), we only presented the questionnaires to new participants.
Finally, all participants answered a few standard LISS evaluation questions about the study, giving them an opportunity to provide written feedback.

@@

Each DDM parameter sub-model was a bi-factor model including the parameter estimates of all tasks as manifest variables.
We estimated a general factor accounting for variance in all manifest variables.
A second ability-specific inhibition factor loaded on the parameter estimates of incongruent trials of the Flanker and Simon tasks.
A third specific attention shifting factor loaded on the parameter estimates of switch trials of the Color-shape, Global-local, and Animacy-size Tasks.
These two factors were allowed to covary.
We also estimated covariances between the conditions of each task.
For each sub-model, we compared the fit of this initial model to a version with a common EF factor loading on incongruent/switch trials of all tasks.
We deemed the second model a better fit when we observed a significant chi squared change test and an AIC value difference \> 10.

For the drift rate model, we interpreted the general factor as basic speed of processing, and the ability-specific factors as reflecting  inhibition and attention-shifting ability (or common EF in the case of the second model).
For the boundary separation and non-decision time models, we interpreted the general factor as general response caution/speed of non-decision processes, and the ability-specific factors as reflecting response caution/speed of non-decision processes specific to incongruent/switch trials.

@@

Finally, we constructed two versions of the final model, one to estimate the association between material deprivation in adulthood and the outcome measures, and the other to estimate the association between threat exposure in adulthood and the outcome measures.
In both models, we regressed the adversity measure on each latent factor, together with the control variables (see the section on confounds for details).
We report indirect effects of control variables in the supplemental materials (section 5).

@@

As preregistered, we first optimized fit in structural equation models for each DDM parameter separately before combining them all into a single model.
For drift rates, we selected the model containing a general processing speed factor (loading on all drift rates) and a common EF factor (loading only on drift rates of incongruent and switch conditions).
An alternative model with a separate inhibition and attention shifting factor did not converge.
The model with a common EF factor provided a good fit to the data (CFI = `r fit_v2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_v2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_v2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_v2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
However, the loadings of the common EF factor were small and in opposing directions, and the latent factor's residual variance was non-significant.
Therefore, we did not estimate associations between adversity and common EF.

@@

```{r}
#| label: Figure3
#| dpi: 600
#| out-width: 15in
#| fig-cap: | 
#|   **Figure 3.** Final measurement model of the Drift Diffusion parameters. Rectangles depict manifest parameter estimates, elipses depict latent factors, unidirectional arrows depict factor loadings, and bidirectional arrows depict correlations. Latent factors with non-significant residual variances are greyed out. *a* = manifest boundary separation, AS = Attention-shifting Task, CS = Color-shape Task, GL = Global-local Task, Fl = Flanker Task, Sm = Simon Task, *t* = manifest non-decision time, *v* = manifest drift rate.
knitr::include_graphics("figures/fig3.png")
```

@@

```{r}
#| tab.id: table4
#| results: markup

table4 |> 
  add_header_row(
    values = " ",
    colwidths = 4
  ) |> # Add a new header row on top. We can use this new row to add the title
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table 4. "), "Fit statistics of all preregistered structural equation models."),
    part = "header"
  ) |>
  border_remove() |> 
  border(i = 1, border.bottom = fp_border_default(), part = "header") |>  
  border(i = 2, border.top = fp_border_default(), part = "header") |> 
  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 15, border.bottom = fp_border_default(), part = "body") |> 
  add_footer_row(
    values = " ",
    colwidths = 4
  ) |> 
  add_footer_row(
    values = " ",
    colwidths = 4
  ) |>  
  flextable::compose(
    i = 1, j = 1, 
    as_paragraph(as_i("Note: "), "for each DDM parameter, measurement model 1 contained a task-general latent factor as well as separate latent factors for inhibition and shifting. Measurement model 2 contained a task-general latent factor as well as a common EF factor. Measurement model 3 contained only a task-general latent factor."), 
    part = "footer"
  ) |> 
  autofit()

@@

Our primary analyses examined the association of exposure to material deprivation and threat in adulthood with: task-general processing speed, task-general response caution, and task-general speed of stimulus encoding and/or response execution.
As the structural equation model analyses reported above showed that EF-specific factors had non-significant variances, we retained them in the model but did not include them as dependent variables in subsequent analyses.

@@

Exposure to material deprivation in adulthood was negatively associated with task-general processing speed ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4). 
People who reported more exposure to deprivation in adulthood processed information more slowly across tasks.
Exposure to deprivation in adulthood was not associated with task-general response caution ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
In addition, none of the effects fell within the region of practical equivalence (all *p*s ≥ `r fit_dep1_eqtests |> pull(eq_pvalue) |> min() |> format_p(pvalue = _, operator = FALSE)`).
The effects in the secondary model (including exposure to threat in adulthood as a confounder) were comparable in size, although the association between material deprivation in adulthood and task-general processing speed did not reach statistical significance.

@@

```{r}
#| label: Figure4
#| fig-width: 6.5
#| fig-height: 5.5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure 4.** Standardized (preregistered) associations between exposure to adversity in adulthood with task-general processing speed, response caution, and non-decision time. The left panel depicts effects of material deprivation and the right panel depicts effects of threat. The top row depicts effects of the primary statistical model (following from our main Directed Acyclic Graph), and the bottom row depicts effects of the secondary statistical model (see Figure 1). For material deprivation in adulthood, the primary statistical model included age, education, and childhood adversity as confounders, whereas the secondary model additionally included threat in adulthood as a confounder. For threat in adulthood, the primary statistial model included age, childhood adversity, and material deprivation in adulthood, whereas the secondary statistical model did not include material deprivation in adulthood as a confounder. The gray area reflects the preregistered area of practical equivalence. Effects depicted with a solid point are practically equivalent, and effects depicted with a hollow point are not practically equivalent. Standard errors represent 95% confidence intervals. The asterisks indicate statistical significance (tested against zero); * *p* < .05, ** *p* < .01, *** *p* < .001.
fig4
```

@@

In the first set of non-preregistered analyses, we analyzed how childhood exposure to threat and material deprivation are associated with task-general and ability-specific cognitive processes.
The rationale for these analyses is that we observed significant indirect associations between childhood adversity and general speed of processing in our preregistered analyses.
Here, we analyzed *direct* effects of childhood threat and material deprivation (i.e., without including exposure to adversity in adulthood).
We controlled for sex (but not age and education) and in the case of threat we also controlled for material deprivation, but not the other way around (similarly to the preregistered analyses).

@@

```{r}
#| label: Figure5
#| fig-width: 6.5
#| fig-height: 3
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure 5.** Standardized (non-preregistered) associations between childhood exposure to adversity with task-general processing speed, response caution, and non-decision time. The left panel depicts effects of material deprivation and the right panel depicts effects of threat. The gray area reflects the preregistered area of practical equivalence. Effects depicted with a solid point are practically equivalent, and effects depicted with a hollow point are not practically equivalent. Standard errors represent 95% confidence intervals. The asterisks indicate statistical significance (tested against zero); * *p* < .05, ** *p* < .01, *** *p* < .001.
fig5
```

@@

In the second set of non-preregistered analyses, we analyzed how adversity exposure during childhood and adulthood were associated with task-specific residual variances.
The rationale for these analyses was that while we did not find coherent latent EF factors, all tasks showed significant residual variance after accounting for general speed of processing.
This implies that, after accounting for general speed of processing, performance is further determined by unique features of individual tasks (i.e., method variance), which may be associated with adversity exposure. 
For this analysis, we were mostly interested in investigating the extent to which task-specific effects reflect EF abilities or processes that are unique to individual tasks, and therefore only focused on drift rates.

@@

Figure 6 shows latent correlations between task-specific drift rates after accounting for task-general processing speed.
The drift rates of repeat and switch conditions within each attention-shifting task correlated moderately to strongly, while all correlations between tasks were low.
This suggests that task-specific drift rates of attention-shifting tasks captured features of the tasks that were not shared between tasks.
In contrast, all correlations between task-specific drift rates of inhibition tasks were low, even among drift rates of the same task.
Thus, task-specific drift rate did provide consistent measures of specific EF abilities that are shared across tasks.

@@

We investigated associations between exposure to two types of adversity in adulthood---material deprivation and threat---with inhibition and attention-shifting ability.
Participants completed two inhibition tasks, three attention shifting tasks, and one basic processing speed task.
First, we used DDM to separate raw performance into three distinct cognitive processes: (1) speed of evidence accumulation (drift rate), (2) response caution (boundary separation), and (3) speed of stimulus encoding and response execution (non-decision time).
Finally, we used structural equation modeling to separate variance in each cognitive process into a task-general factor (shared across all tasks) and ability-specific factors.

@@

People with more exposure to threat in adulthood---but not material deprivation---showed lower general processing speed.
This is consistent with predictions from deficit frameworks [@sheridan_2014; @tucker_drob_2013].
After accounting for general processing speed, there was no remaining variance that could be attributed to either inhibition ability or attention-shifting ability.
Additionally, the associations between material deprivation and threat in adulthood with both response caution and non-decision time were neither significantly different from zero, nor practically equivalent to zero.
In other words, from these data, we were neither able to conclude that people with more exposure to adversity differed in their level of response caution and non-decision time, nor that these processes were similar to people with lower levels of adversity exposure.

We found that variance in EF task performance was mostly explained by a task-general factor, not by ability-specific factors.
This finding aligns with previous research applying DDM to EF tasks. 
Three recent studies converged on the similar conclusion that performance across different EF tasks is best explained by a task-general drift rate factor [@frischkorn_2019; @hedge_2022; @loffler_2024; @weigard_2021].
One study found that the task-general drift rate factor fully reflected basic processing speed, rather than common EF processes [@loffler_2024].
Similarly, another study including several inhibition tasks found that shared variance was mostly explained by basic processing speed and strategy differences, not by inhibition ability [@hedge_2022].
It is therefore not surprising that we were unable to estimate ability-specific EF factors after accounting for general processing.
These issues tie into several other criticisms of common EF tasks in the cognitive psychology literature---such as low test-retest reliability---which should be considered when using these tasks to test associations with adversity [@vonbastian_2020].
